Append-only Datastore

### 2020-06-01 paper/paper.tex

\section{Design}
% TODO: Add architecture diagram.
Nodes in the append-only datastore are classified in a simple hierarchy as seen
in TODO(add the architecture diagram). A single \textit{leader} directs multiple
\textit{followers}. Clients connnect to and communicate with a nearby follower,
likely running in the same datacenter, to minimize latency. Our design tolerates
arbitary partitions between nodes and ensures an eventually consistent view of
data.

@@

Followers use the \texttt{sync} request to get data they are missing,
effectively ``patching the holes" in their local datastore. This ensures the
eventually consistent property of the system. \texttt{sync} requests also signal
what data the sender has to the recipient, enabling the recipeint to also patch
holes in its local datastore.

@@

\begin{enumerate}
	\item Write the append to local storage along with the index \texttt{i}
		the follower believes applies to the append.
	\item Respond to the client that the request was successful.
	\item Concurrently with (2), send an \texttt{update} to the leader that
		an append occured on key \texttt{k} at index \texttt{i}.
	\item The leader responds with \texttt{SUCCESS}.
	\item The follower updates local disk to indicate that the append has
		been commited at index \texttt{i} by the leader.
\end{enumerate}

The data for key \texttt{k} has simply benn appended to, changing from $\{d_a,
d_b, d_c, ..., d_{i-1}\}$ to $\{d_a, d_b, d_c, ..., d_{i-1}, d_i\}$, where $a < b <
c < i$. Note that there may be data ``holes" between $a$, $b$, $c$, and $i$ not
yet filled in by either a \texttt{get} or \texttt{sync}.

\subsubsection{Conccurent Updates}
When concurrent updates occur:

% TODO: There's some ugly mixing of texttt and math notation here, particularly
% around i'.
\begin{enumerate}
	\item The same as steps 1-3 above.
	\item The leader responds with \texttt{NEEDS\_UPDATE}.
	\item The follower updates local disk to indicate that the append has
		been commited at index \texttt{i`} by the leader.
	\item The follower sends a \texttt{sync} request to the follower that
		appended the data at index \texttt{i`-1}, as indicated in the
		leader's response.
	\item The follower receives and stores on local disk the data it was
		missing, including what the leader deemed to be the ``winner" of
		index \texttt{i}.
\end{enumerate}

@@

The system continues to function when the leader is partitioned, but followers
and thus clients do not receive updates from other followers. As long as the
partition is eventually healed, the system will propgate information correctly
and self-heal. If for some reason there is a permanent partition, it must be
manually worked around by changing the cluster configuration (i.e. the set of
nodes in the system).

@@

\begin{itemize}
	\item Can read their own appends immiedately.
	\item Never see data reordered relatively.
	\item Communicate only with a fast, nearby follower, and no other nodes
		in the system.
	\item Can always read and write, even when their follower is partitioned
		from the rest of the system.
\end{itemize}

@@

\section{Future Work}
We could greatly increase throughput by providing a client library (rather than
a raw GRPC interface) that is aware of the version or latest verions it holds.
This would enable followers to selectively return only the few missing pieces of
data a client is missing rather than all the data.

To increase fault tolerance, we have discussed making each follower and the
leader their own small cluster of concensus nodes. This would greatly increase
fault tolerance and, while it may complicate the system, would not complicate
clients or the protocols via which nodes communicate.

>>>>

\section{Design}
% TODO: Add architecture diagram.
Nodes in the append-only datastore are classified in a simple hierarchy as seen
in TODO(add the architecture diagram). A single \textit{leader} directs multiple
\textit{followers}. Clients connect to and communicate with a nearby follower,
likely running in the same datacenter, to minimize latency. Our design tolerates
arbitrary partitions between nodes and ensures an eventually consistent view of
data.

@@

Followers use the \texttt{sync} request to get data they are missing,
effectively ``patching the holes" in their local datastore. This ensures the
eventually consistent property of the system. \texttt{sync} requests also signal
what data the sender has to the recipient, enabling the recipient to also patch
holes in its local datastore.

@@

\begin{enumerate}
	\item Write the append to local storage along with the index \texttt{i}
		the follower believes applies to the append.
	\item Respond to the client that the request was successful.
	\item Concurrently with (2), send an \texttt{update} to the leader that
		an append occurred on key \texttt{k} at index \texttt{i}.
	\item The leader responds with \texttt{SUCCESS}.
	\item The follower updates local disk to indicate that the append has
		been committed at index \texttt{i} by the leader.
\end{enumerate}

The data for key \texttt{k} has simply been appended to, changing from $\{d_a,
d_b, d_c, ..., d_{i-1}\}$ to $\{d_a, d_b, d_c, ..., d_{i-1}, d_i\}$, where $a < b <
c < i$. Note that there may be data ``holes" between $a$, $b$, $c$, and $i$ not
yet filled in by either a \texttt{get} or \texttt{sync}.

\subsubsection{Concurrent Updates}
When concurrent updates occur:

% TODO: There's some ugly mixing of texttt and math notation here, particularly
% around i'.
\begin{enumerate}
	\item The same as steps 1-3 above.
	\item The leader responds with \texttt{NEEDS\_UPDATE}.
	\item The follower updates local disk to indicate that the append has
		been committed at index \texttt{i`} by the leader.
	\item The follower sends a \texttt{sync} request to the follower that
		appended the data at index \texttt{i`-1}, as indicated in the
		leader's response.
	\item The follower receives and stores on local disk the data it was
		missing, including what the leader deemed to be the ``winner" of
		index \texttt{i}.
\end{enumerate}

@@

The system continues to function when the leader is partitioned, but followers
and thus clients do not receive updates from other followers. As long as the
partition is eventually healed, the system will propagate information correctly
and self-heal. If for some reason there is a permanent partition, it must be
manually worked around by changing the cluster configuration (i.e. the set of
nodes in the system).

@@

\begin{itemize}
	\item Can read their own appends immediately.
	\item Never see data reordered relatively.
	\item Communicate only with a fast, nearby follower, and no other nodes
		in the system.
	\item Can always read and write, even when their follower is partitioned
		from the rest of the system.
\end{itemize}

@@

\section{Future Work}
We could greatly increase throughput by providing a client library (rather than
a raw GRPC interface) that is aware of the indeces or latest index it holds.
This would enable followers to selectively return only the few missing pieces of
data a client is missing rather than all the data.

To increase fault tolerance, we have discussed making each follower and the
leader their own small cluster of consensus nodes. This would greatly increase
fault tolerance and, while it may complicate the system, would not complicate
clients or the protocols via which nodes communicate.
